---
title: "Bayesian Linear Regression"
author: "Matilde Nielsen"
date: "`r Sys.Date()`"
output:
  md_document:
    variant: markdown_github
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(blm)
library(MASS)
```

## Bayesian Linear Regresssion

This packages was built for the class "Datascience II - Software Development and Testing" at Aarhus University, January 2017. It implements methods to construct ´blm´ objects, calculating posterior distributions and predicting responses from data.

The definition of a linear model is any model that takes on the form

![equation](http://www.sciweavers.org/tex2img.php?eq=Y%3Dw_0%2B%5Csum_%7Bi%3D1%7D%5Enw_if%28x_i%29&bc=White&fc=Black&im=jpg&fs=12&ff=arev&edit=0)

for random variables $x_i \in [1;n]$ where $w_i$ respresents the weight associated with the term $f(x_i)$. In general, the function $f(x_i)$ can be any function, but this package only deals with cases where $f(x_i)$ is a linear function. It is also assumed that the data is normally distributed.

The linear model must be built using a training dataset that contains one or more explanatory variables and a response variable. The `blm` constructor requires a prior distribution of weights (i.e. their mean and their variance). This can be any distribution, but the `blm` package provides a function `make_prior` that can construct this for you, based on a given prior precision, $\alpha$. This will output a covariance matrix of the form

![equation](http://www.sciweavers.org/tex2img.php?eq=%5Csigma_%7Bij%7D%3D%5Cbegin%7Bcases%7D1%2F%5Calpha%2C%26%5Ctext%7Bif%20i%3Dj%7D%5C%5C0%2C%26%5Ctext%7Botherwise%7D%5Cend%7Bcases%7D&bc=White&fc=Black&im=jpg&fs=12&ff=arev&edit=0)

and a mean of 0 for all $w_i$.

Given a model, a prior distribution, a posterior precision $\beta$, and some data, `blm` then calculates a posterior distribution of the weights. This is a normal distribution with $w_i \sim N(m_{x,y},\sigma_{x,y})$ where 

$$\mathbf{m}_{x,y}=\beta \:\mathbf{S}_{x,y} {\phi_X}^T \mathbf{y},$$

$$\sigma_{x,y}^{-1}=\alpha I + \beta\:{\phi_X}^T\phi_X$$

$\phi_X$ is the model matrix where the first column has $1$ in every row and each explanatory variable has it's own column.
Below is shown the results for a simulated data set.
```{r simulateTraining}
#Simulate training data:
set.seed(1)
beta = 0.2; alpha = 1
w0 = 0.2; w1 = 1.2; w2 = 0.5
d = data.frame(x=rnorm(500), z=rnorm(500))
d$y = rnorm(500, w0+w1*d$x+w2*d$z, 1/beta)
head(d)
```
```{r myBlm}
#The model:
m = y ~ x+z

#The prior:
myPrior = make_prior(m, alpha)

#The bayesian linear model:
myBlm = blm(m, myPrior, beta, d)

#Predictions based on the training data:
myFit = fitted(myBlm)
head(myFit)

summary(myBlm)
plot(myBlm)
```

Now, we can use the bayesian linear model to predict the response for new data:
```{r newData}
#New data:
set.seed(2)
d2 = data.frame(x=rnorm(20), z=rnorm(20))
d2

#Predict response variables for new data using the blm:
myPredict = predict(myBlm, d2)

```

We can compare these results with a regular linear model (using `lm` from package `stats`):

```{r lm}
#Creating a linear model:
linReg = lm(m,d)
#Predicting the repsonse using linear model:
linRegPredict = predict(linReg, d2)
plot(linReg, which=c(1,2))

#Comparing blm vs lm predictions:
blmVSlm = lm(linRegPredict ~ myPredict$Prediction)
summary(blmVSlm)
plot(linRegPredict, myPredict$Prediction, main="blm vs lm predictions", xlab="lm predictions", ylab="blm predictions")
abline(a=blmVSlm$coefficients[1], b=blmVSlm$coefficients[2], col="red")

#compare residuals:
blmRes = residuals(myBlm)
lmRes = linReg$residuals

resCompare = lm(abs(lmRes) ~ abs(blmRes[,1]))
summary(resCompare)

#Plotting the absolute residuals:
plot(abs(lmRes), abs(blmRes[,1]), main="Comparison of residuals from blm and lm", xlab="Residuals from lm", ylab="Residuals from blm")
abline(a=resCompare$coefficients[1], resCompare$coefficients[2], col="red")
```

This shows that the two types of linear regressions are highly correlated (with a slope very close to 1 and and intercept not significantly different from 0) for a data set of this size.



